**The Mint Scale: A Fresh Validation of the Multimodal Interoception Questionnaire and Comparison to the MAIA, BPQ and IAS**

### 2025-08-12 paper/manuscript.qmd
push

---
title: "**The Mint Scale: A Fresh Validation of the Multimodal Interoception Questionnaire and Comparison to the MAIA, BPQ and IAS**"
shorttitle: "Mint Validation"
author:
  - name: Dominique Makowski
    corresponding: true
    orcid: 0000-0001-5375-9967
    email: D.Makowski@sussex.ac.uk
    url: https://realitybending.github.io/
    roles:
      - Conceptualization
      - Data curation
      - Formal Analysis
      - Funding acquisition
      - Investigation
      - Methodology
      - Project administration
      - Resources
      - Software
      - Supervision
      - Validation
      - Visualization
      - Writing – original draft
    affiliations:
      - id: "id1"
        name: "University of Sussex"
        department: "School of Psychology"
        city: "Brighton"
        country: "UK"
        postal-code: "BN1 9RH"
      - name: "University of Sussex"
        department:  "Sussex Centre for Consciousness Science"
  - name: Ana Neves  
    orcid: 0009-0006-0020-7599
    role:
      - Investigation
      - Methodology
      - Project administration
      - Supervision
      - Writing – original draft
      - Writing – review & editing
    affiliations:
      - ref: "id1"
  - name: Emma Benn 
    role:
      - Investigation
    affiliations:
      - ref: "id1"
  - name: Maisie Bennett
    role:
      - Investigation
    affiliations:
      - ref: "id1"
  - name: Giulia Poreiro
    orcid: 0000-0002-2343-5109
    role:
      - Investigation
      - Methodology
      - Writing – original draft
      - Writing – review & editing
    affiliations:
      - ref: "id1"
author-note:
  status-changes: 
    affiliation-change: ~
    deceased: ~
  disclosures:
    gratitude: |
      ::: {.callout-note icon=false appearance="simple"}
      This preprint is a non-peer-reviewed work from the [**Reality Bending Lab**](https://realitybending.github.io/).
      ![](https://realitybending.github.io/media/ReBeL_LogoOnly_hu11484441381606756729.png){width=20% fig-align="center"}
      :::
    study-registration: ~
    data-sharing: ~
    related-report: ~
    conflict-of-interest: ~
    financial-support: ~
    authorship-agreements: ~
abstract: |
  TO DO.
#   **Significance Statement**. TODO.
keywords: "Interoception questionnaire, interoceptive accuracy scale, MAIA, Mint Validation, Body Awareness"
numbered-lines: true
floatsintext: true
mask: false
bibliography: bibliography.bib
format:
  apaquarto-pdf:
    documentmode: jou
    keep-tex: true
  apaquarto-docx: default
---

@@

Main issues in existing questionnaires:

- Either heavily based on theories (e.g., focusing on specific "alleged" dimensions), despite shaky evidence for said-theories
- Do not control for context (which leads to variability in interpretation and occurence)
- Often quite narrow in the modalities covered (or scattered, like the IAS)

@@

TODO: write intro.

- Goal of study 1: to generate a lot of items, analyze its structure and reduce them to a balanced set of items.

@@

After removing redundant items using UVA, we iteratively fitted hierarchical EGA models (which additionally estimates higher-order "meta" clusters) using "glasso" [@friedman2008sparse] and the "Leiden" algorithm [@traag2019louvain] for community detection, refining the item pool at each step. We started by removing items with a low (< 80%) cluster stability (i.e., volatile items which jump between clusters across bootstrapped samples), followed by odd items belonging to no clusters or pairs of items (i.e., we keep items belonging to clusters of more than 2 items). Finally, for each lower-level cluster, we selected the 3 items with the highest node centrality (i.e., the highest loading in the cluster).

@@

- Possible (although not at all significant) bias consistent across indices in favour of a greater similarity with the modality-grouping condition. Futher research with more data (here we only had ~200 observation per condition) might be able to yield more conclusive results. In general, testing the presentation mode is a novel approach which should be more widely applied to psychometric validations.
- We grouped some of the dimensions under "Awareness", despite being similar to items included in the IAS typically conceptualized as "Accuracy". However, we believe that subjective scales cannot be pure measures of accuracy, as they are inherently subjective and thus reflect a degree of awareness (of one's accurate noticing or sensation).
- Olfactory compensation: pseudo-interoceptive pathway whereby people with interoceptive deficits would compensate by gathering evidence about their internal state externally (e.g., via smell).
- The metacluster "Sensitivity" seems to be grouping iterms related to visceropcetion and (hyper)sensitivity (Derm).

@@

- Goal of study 2: to validate the Mint scale against other interoception scales.

@@

Participants were asked to select somatic symptoms or conditions they experienced from a list of 36 options.
To facilitate mechanistic interpretation and reduce redundancy, answers were grouped into four non-overlapping clusters based on shared physiological pathways and known etiological mechanisms. The *Afferent Sensitivity* cluster included conditions associated with heightened interoceptive awareness and neurogenic excitability, such as migraine, neuropathy, dizziness, nausea, muscle tension, epilepsy, and frequent urination. The *Central Sensitization* cluster comprised syndromes characterized by chronic pain and fatigue, likely reflecting central amplification of sensory signals and HPA-axis dysregulation, such as fibromyalgia, chronic fatigue syndrome, chronic pain, back pain, pelvic pain, irritable bowel syndrome (IBS). The *Autonomic Dysfunction* cluster captured disorders linked to dysregulation of the autonomic and cardiopulmonary systems, including joint hypermobility, cardiac arrhythmia, chest pain, shortness of breath, hypo-/hypertension, sleep apnea, chronic obstructive pulmonary disease (COPD), and chronic bronchitis. Finally, the *Immune-Inflammatory* cluster encompassed conditions associated with immune dysregulation, barrier dysfunction, and gut-brain axis disturbance, such as eczema, psoriasis, skin rashes, asthma, celiac disease, gluten and lactose sensitivity, inflammatory bowel diseases (Crohn’s disease and ulcerative colitis), gastroesophageal reflux disease (GERD), multiple sclerosis, and Sjögren’s syndrome. 
<!-- Each symptom was assigned to a single cluster to preserve orthogonality and support subsequent analyses linking symptom profiles to interoceptive processes. -->
We scored each cluster as a binary variable based on whether the participant selected at least one symptom from that cluster.

@@

- Structure: Low stability of *Olfa* as belonging in the deficit group. More studies are needed to investigate its place and cluster invariance: it is possible that its association might depend on other factors or categories.
- Main contender is MAIA. But its most sensitive dimensions seem to be the least related to interoception, which is dangerous as it might lead to findings with associations misattributed to "Interoception" (just because it correlate with some MAIA dimensions), but it might have nothing to do with interoception per se and more with emotion regulation. 
- Basically the pattern is clear: everything high-level, metacognitive and highly subjective related to wellbeing and general affective state is the most strongly related to the MAIA through some of its Emotion regulation dimensions. All the other outcomes actually related to interoceptive processes are most strongly predicted by the Mint.

@@

- Future directions: needs to test relationship with physiological measures of interoception.
- Also, the structure might still require a bit of deeper investigation, in particular with some of its facets (*Olfa* and *Derm*) which might display different associations depending on the context or the population.

@@

We would like to thank the dissertation students from the University of Sussex for their help in data collection.
DM would also like to thank ... for the motivation provided to write this paper.

>>>>

---
title: "**The Mint Scale: A Fresh Validation of the Multimodal Interoception Questionnaire and Comparison to the MAIA, BPQ and IAS**"
shorttitle: "Mint Validation"
author:
  - name: Dominique Makowski
    corresponding: true
    orcid: 0000-0001-5375-9967
    email: D.Makowski@sussex.ac.uk
    url: https://realitybending.github.io/
    roles:
      - Conceptualization
      - Data curation
      - Formal Analysis
      - Funding acquisition
      - Investigation
      - Methodology
      - Project administration
      - Resources
      - Software
      - Supervision
      - Validation
      - Visualization
      - Writing – original draft
    affiliations:
      - id: "id1"
        name: "University of Sussex"
        department: "School of Psychology"
        city: "Brighton"
        country: "UK"
        postal-code: "BN1 9RH"
      - name: "University of Sussex"
        department:  "Sussex Centre for Consciousness Science"
  - name: Ana Neves  
    orcid: 0009-0006-0020-7599
    role:
      - Investigation
      - Methodology
      - Project administration
      - Supervision
      - Writing – original draft
      - Writing – review & editing
    affiliations:
      - ref: "id1"
  - name: Emma Benn 
    role:
      - Investigation
    affiliations:
      - ref: "id1"
  - name: Maisie Bennett
    role:
      - Investigation
    affiliations:
      - ref: "id1"
  - name: Giulia Poreiro
    orcid: 0000-0002-2343-5109
    role:
      - Investigation
      - Methodology
      - Writing – original draft
      - Writing – review & editing
    affiliations:
      - ref: "id1"
author-note:
  status-changes: 
    affiliation-change: ~
    deceased: ~
  disclosures:
    gratitude: |
      ::: {.callout-note icon=false appearance="simple"}
      This preprint is a non-peer-reviewed work from the [**Reality Bending Lab**](https://realitybending.github.io/).
      ![](https://realitybending.github.io/media/ReBeL_LogoOnly_hu11484441381606756729.png){width=20% fig-align="center"}
      :::
    study-registration: ~
    data-sharing: ~
    related-report: ~
    conflict-of-interest: ~
    financial-support: ~
    authorship-agreements: ~
abstract: |
  TO DO.
#   **Significance Statement**. TODO.
keywords: "Interoception questionnaire, interoceptive accuracy scale, MAIA, Mint Validation, Body Awareness, Health"
numbered-lines: false
floatsintext: true
mask: false
bibliography: bibliography.bib
format:
  apaquarto-pdf:
    documentmode: jou
    keep-tex: true
  apaquarto-docx: default
---

@@

Interoception - the sensing and interpretation of the body's internal physiological signals - plays a fundamental role in processes ranging from homeostatic regulation to emotion, decision-making, and mental health [@craig2002how; @critchley2017interoception]. Despite its growing recognition, the field is hindered by inconsistent definitions and fragmented measurement approaches [@khalsa2018interoception; @desmedt2025discrepancies; @desmedt2022measures]. Available tools include "objective" physiological tasks (e.g., heartbeat counting or tracking) and "subjective" self-reports, and encompass both "explicit" paradigms that direct participants' attention and consciousness to internal sensations and "implicit" paradigms that do not (e.g., heartbeat evoked potentials recorded during rest). These assessments are often targeted at only one of several interoceptive modalities (e.g., cardioception, respiroception, gastroception), and are mapped onto theoretical dimensions [e.g., accuracy, sensitivity, awareness, @garfinkel2015knowing] within a rapidly evolving theoretical scaffolding. However, no consensus exists on a gold-standard measure, and each approach has distinct limitations: physiological tasks often require specialised equipment and expertise, are time-consuming, and are challenging to administer at scale, whereas questionnaires are constrained by their inherent subjectivity. Crucially, the psychometric quality of many of these measures is questionable [e.g., see criticism about the heartbeat counting task, @zamariola2018interoceptive], yielding low inter-task correlations [@murphy2018classifying], raising serious concerns about their reliability and validity (e.g., measures presented as measuring one theoretical dimension but likely measuring something else). This measurement conundrum hinders progress in comparing, replicating, and interpreting interoception-related findings, underscoring the need to develop improved tools that align with modern psychometric standards.

<!-- There are many questionnaire measures – 14 of which are commonly used and 5 (BPQ, MAIA, BAQ, SAQ, and PBCS) which cover 86.5% of all citations of interoceptive sensibility questionnaires – Desmedt et al., 2022 -->
The present work focuses on self-report questionnaires, which offer a scalable and accessible tool complementing physiological tasks. Among the numerous questionnaires available, a handful seem to dominate the literature [@desmedt2022measures], including the Body Perception Questionnaire [BPQ, @kolacz2018body], the Multidimensional Assessment of Interoceptive Awareness [MAIA, @mehling2018multidimensional], and the Interoceptive Accuracy Scale [IAS, @murphy2018classifying]. These tools were developed in different eras of a rapidly evolving field, each with its own objectives and philosophies. The BPQ, one of the earliest measures, emphasizes the assessment of body awareness and autonomic nervous system (ANS) reactivity, focusing on subjective experiences of physiological stress reactions in organs innervated by the ANS, such as those triggered by stress or trauma [@porges1993body]. The MAIA incorporated ideas from the then-booming emotion regulation and mindfulness fields, including subscales assessing adaptive coping strategies and metacognitive insights about the body. By contrast, the IAS was recently crafted to sidestep such high-level - possibly tangential - constructs and instead capture subjective reports of concrete bodily events (e.g., coughing, sneezing) as proxies for interoceptive accuracy.

These differing philosophies have practical consequences, leading to poor convergence between scales that are assumed to measure the same construct [@vig2022questionnaires] - a classic example of the "jingle fallacy" where the shared label of "interoception" can mask underlying measurand differences, which in turn hinders replication and theoretical integration. Recent work has been focused on aligning subjective measures and facets with theoretical models, debating whether they measure interoceptive "sensitivity" or other dimensions (accuracy, attention, deficits). However, this focus on a theory-driven approach might be premature when theoretical models are in flux, and when a topographic mapping to self-report dimensions is not guaranteed (e.g. distinctions that are relevant at a neuroanatomical or cognitive level might not necessarily translate into the same structural landscape of subjective experiences). More fundamentally, existing questionnaires suffer from two major structural blind spots. First, they are often narrow or unsystematic in their coverage of bodily *modalities*, with a heavy bias towards cardio-respiratory sensations and a neglect of others like gastric, urogenital, or thermoregulatory signals. Second, and perhaps most critically, they fail to control for the *context* in which an interoceptive experience occurs, introducing significant variability in how participants interpret and respond to items. For example, an item about "feeling your heart beat" is ambiguous; its frequency, experience and its interpretation differ vastly depending on whether it occurs during physical exercise, a moment of anxiety, or quiet rest (e.g., someone who exercises regularly, or is anxious, might more easily recall an instance of occurrence of feeling their heart beating and respond based on that contextual saliency).

To address these gaps, we introduce the Multimodal Interoception Questionnaire (Mint). The overarching aim of this work is to develop and validate a new questionnaire that systematically addresses these limitations. In Study 1, we detail the bottom-up development of the Mint, using a "modality-by-context" framework to generate a comprehensive item pool, which is then refined using cutting-edge psychometric network analysis. In Study 2, we validate the final version of the Mint against the established interoception questionnaires, demonstrating its robust structure and predictive validity for a range of clinically-relevant outcomes.

@@

The primary goal of Study 1 was to develop a new interoception questionnaire by generating a comprehensive item pool and refining it into a psychometrically robust scale. To overcome the limitations of existing measures, we adopted a bottom-up, theoretically agnostic approach. Instead of imposing a pre-defined theoretical structure, we focused on two tangible and systematically controllable aspects of interoceptive experience: the *modality* of the sensation and the *context* in which it occurs. The objective was not to create *yet another scale* measuring an alleged distinct construct or variation, but rather to synthesize existing ideas and frameworks into a comprehensive and modern tool that captures the richness of interoceptive experiences and encompasses the most relevant facets of interoception.

<!-- Give examples of items for each part of the framework here so that the method makes more sense? -->
Our item generation process was guided by a "modality-by-context" grid designed to overcome the limitations of existing scales. For *modalities*, we expanded beyond the typical cardio-respiratory focus to include a broader range of bodily systems: cardiac, respiratory, gastric, genital, bladder & colon, and skin & temperature [the latter of which being increasingly recognized as a key interoceptive channel, @crucianelli2023role]. We also included a "general state" category to capture the integrated sense of the body's overall physiological condition [@craig2002how; @craig2003interoception]. For *contexts*, we designed facets to capture specific situational contexts like negative (anxious) and positive (sexual) arousal, which reflect distinct bodily states with differing interoceptive signatures, as well as other experiential qualities tapping into elaboration and threshold of perception, such as sensitivity, accuracy, and confusion. To maintain continuity with established work while making targeted improvements, we purposefully adapted items from existing questionnaires (e.g., IAS-like items related to observable behaviors). This systematic framework yielded a diverse pool of 120 initial items, ensuring comprehensive coverage while minimizing the ambiguity inherent in context-free questions.

The final step involved a data-driven reduction of this item pool. We employed cutting-edge psychometric network-based techniques to identify and remove redundant items and to uncover the most stable and coherent dimensional structure. Additionally, we designed the study so that we could explore an often overlooked psychometric quality: the invariance of the structure to the presentation format (how items are grouped together on the screen, whether by modality, facet, or randomly) influences the resulting psychometric structure. This rigorous, multi-stage process was designed to produce a subset of items with a clear, stable, and theoretically rich hierarchical structure, ready for formal validation.

@@

After removing redundant items using UVA, we iteratively fitted hierarchical EGA models (which additionally estimate higher-order "meta" clusters) using "glasso" [@friedman2008sparse] and the "Leiden" algorithm [@traag2019louvain] for community detection, refining the item pool at each step. We started by removing items with a low (< 80%) cluster stability (i.e., volatile items which jump between clusters across bootstrapped samples), followed by odd items belonging to no clusters or pairs of items (i.e., we keep items belonging to clusters of more than 2 items). Finally, for each lower-level cluster, we selected the 3 items with the highest node centrality (i.e., the highest loading in the cluster).

@@

<!-- summary -->
The primary aim of Study 1 was to develop and refine a large, systematically generated pool of interoceptive items into a psychometrically sound and structurally coherent questionnaire. Starting with 120 items derived from a modality-by-context grid, we employed a multi-step, data-driven reduction process using cutting-edge structure analysis methods. This rigorous procedure yielded a final, balanced set of 45 items, organized within a stable hierarchical structure comprising three higher-order dimensions (Interoceptive Deficits, Interoceptive Awareness, Interoceptive Sensitivity) and 12 distinct lower-level facets.

<!-- Effect of item presentation -->
One innovation of this study was to test the effect of item presentation by comparing structures across different grouping conditions. While the analysis revealed no significant effect, a consistent (though non-significant) trend suggested that grouping items by modality may be slightly more conducive to revealing the underlying psychometric structure than grouping by facet or randomly. This finding, while preliminary due to the sample size per condition ($N \pm 200$), highlights a potentially important and under-investigated aspect of questionnaire design that warrants future research.
<!-- The final reduction from 54 to 45 items was a consensus-based decision to remove three clusters for reasons of conceptual clarity and parsimony, a necessary step to ensure the final scale's usability and content validity. -->

The resulting hierarchical structure offers novel yet interpretable insights into self-reported interoception, with flexible degrees of granularity. Notably, we categorised several facets (typically related to objective external behaviours, like coughing or sneezing) under the label "Awareness", that might traditionally be named "Accuracy" in other scales. Our rationale is that subjective self-report, by its very nature, cannot be a pure measure of objective accuracy, but at best a proxy of an individual's subjective certainty or awareness of their interoceptive processes. Furthermore, our bottom-up approach revealed intriguing facets, such as "Olfactory Compensation". This finding provide further evidence for a pseudo-interoceptive pathway, whereby individuals with potentially poorer access to internal signals may compensate by relying on external cues (e.g., odours) to infer their bodily state. The "Sensitivity" metacluster also grouped core visceroceptive facets (e.g., "Cardioception", "Gastroception") with "Dermal Hypersensitivity", pointing to a potential common dimension of heightened perceptual sensitivity to bodily sensations.

In summary, Study 1 successfully leveraged a systematic item generation process and modern analytical techniques to produce a 45-item questionnaire with a clear, stable, and theoretically rich hierarchical structure. The next crucial step is to formally validate this new instrument by examining its psychometric properties and its relationship with existing measures in a new, independent sample, which is the objective of Study 2.

@@

Following the generation and selection of a set of 45 interoceptive items, the next objective is to formally validate this new instrument in a large, independent sample, a crucial step for establishing the psychometric soundness and unique contribution of the Mint scale to the field of interoception research. To this end, we sought to 1) re-assess the structure of the shorter version and 2) establish its convergent validity by examining the correlations with other established and popular interoception questionnaires (namely the MAIA-2, BPQ, and IAS), as well as other related measures (e.g., alexithymia). Finally, we assessed its criterion validity by directly comparing the predictive power of the interoceptive scales across a broad range of outcomes, including emotional, cognitive, and clinical variables.

@@

Participants were asked to select somatic symptoms or conditions they experienced from a list of 36 options.
To facilitate mechanistic interpretation and reduce redundancy, answers were grouped into four non-overlapping clusters based on shared physiological pathways and known etiological mechanisms. The *Afferent Sensitivity* cluster included conditions associated with heightened interoceptive awareness and neurogenic excitability, such as migraine, neuropathy, dizziness, nausea, muscle tension, epilepsy, and frequent urination. The *Central Sensitization* cluster comprised syndromes characterized by chronic pain and fatigue, likely reflecting central amplification of sensory signals and HPA-axis dysregulation, such as fibromyalgia, chronic fatigue syndrome, chronic pain, back pain, pelvic pain, irritable bowel syndrome (IBS). The *Autonomic Dysfunction* cluster captured disorders linked to dysregulation of the autonomic and cardiopulmonary systems, including joint hypermobility, cardiac arrhythmia, chest pain, shortness of breath, hypo-/hypertension, sleep apnea, chronic obstructive pulmonary disease (COPD), and chronic bronchitis. Finally, the *Immune-Inflammatory* cluster encompassed conditions associated with immune dysregulation, barrier dysfunction, and gut-brain axis disturbance, such as eczema, psoriasis, skin rashes, asthma, celiac disease, gluten and lactose sensitivity, inflammatory bowel diseases (Crohn's disease and ulcerative colitis), gastroesophageal reflux disease (GERD), multiple sclerosis, and Sjögren's syndrome. 
<!-- Each symptom was assigned to a single cluster to preserve orthogonality and support subsequent analyses linking symptom profiles to interoceptive processes. -->
We scored each cluster as a binary variable based on whether the participant selected at least one symptom from that cluster.

@@

The primary goal of Study 2 was to validate the items selected in Study 1. This involved re-assessing its structure in a new, large sample, and establishing its convergent and criterion validity against other established interoception questionnaires. The analysis led to a final, refined 33-item Mint scale with a stable and theoretically coherent hierarchical structure, which demonstrated substantial predictive power across a range of outcomes.

Re-validating the structure after item selection is a crucial step in questionnaire development. When a large item pool is reduced, the context changes, which can alter the perceived meaning and relationships between the remaining items. For instance, a questionnaire with many items suggesting abnormality or pathology might colour other neutral items with a negative connotation. While the original structure was mostly replicated, the re-validation allowed for a clarification of some of the higher-order dimensions (for instance, the "Sensitivity" metacluster became more clearly defined around core interoceptive clusters). The final structure, comprising 11 facets and 3 metaclusters, showed excellent structural consistency and stability, providing a solid foundation for the scale. However, the *Olfa* (Olfactory Compensation) cluster showed lower stability in its placement within the *Deficit* metacluster. This suggests its role may be more complex, and future studies are needed to investigate its place and invariance across different populations and contexts.

The criterion validity analysis revealed a clear and compelling pattern of results. The Mint scale, particularly its *Deficit* metacluster, consistently outperformed other scales in predicting outcomes directly related to alterations in bodily experience and self-regulation, such as alexithymia, dissociative experiences, ADHD, autism, and various somatic symptoms. The main contender, the MAIA-2, showed an advantage primarily for outcomes related to general well-being, such as life satisfaction, depression, and anxiety. However, this advantage was almost entirely driven by dimensions highly related to emotion regulation abilities, namely *Trusting* and *Not-Worrying*, whose inclusion as core interoceptive facets has been questioned [@ferentzi2021examining]. When these dimensions were removed (in the *iMAIA* version), the MAIA-2's predictive power for these outcomes dropped considerably, often below that of the Mint.

This highlights a crucial issue in the field: the potential for misattributing findings to "interoception" when they may be driven by broader emotion regulation capacities [@murphy2020testing]. While valuable, the MAIA-2's most predictive subscales appear less related to core interoceptive processing and more to metacognitive beliefs about one's body and emotional coping. The Mint, in contrast, appears to tap more directly into distinct facets of interoceptive experience, from visceral sensitivity to perceived deficits. The pattern is clear: for high-level, metacognitive aspects of affective state and well-being, the MAIA-2's emotion regulation facets are highly relevant. For outcomes tied more specifically to bodily processes, clinical conditions characterized by sensory dysregulation, and nuanced interoceptive experiences, the Mint offers superior predictive validity and specificity.

@@

Across two studies, we developed and validated the Multimodal Interoception Questionnaire (Mint), a new 33-item self-report measure of interoception. Our approach addressed key limitations of existing scales by starting with a systematically generated, broad item pool that explicitly controlled for context and modality, and then using a data-driven reduction process. The result is a psychometrically robust instrument with a stable, hierarchical structure comprising three higher-order metaclusters, namely *Interoceptive Deficit*, *Interoceptive Awareness*, and *Visceroception*, and 11 distinct lower-level facets. The validation study demonstrated that the Mint not only captures variance present in other popular scales (MAIA-2, BPQ and IAS) but also offers superior predictive power for a range of clinical and cognitive outcomes, particularly those tied to altered bodily perception.

A key contribution of the Mint is its potential ability to disentangle core interoceptive experiences from more general emotion regulation capacities. Our findings show that while the MAIA-2 is a potent predictor of general well-being, this is largely driven by its *Trusting* and *Not-Worrying* subscales, which heavily tap into emotion regulation skills and coping strategies. The Mint, conversely, excelled at predicting conditions like alexithymia, ADHD, autism, and somatic symptoms, where alterations in bodily sensing are a core feature. This suggests the Mint provides a more specific measure of interoceptive processing, offering researchers a finer-grained tool to probe the role of bodily awareness in health and disease.

Despite these strengths, the present work has limitations that open avenues for future research. First, as a self-report measure, the Mint captures subjective beliefs about interoceptive abilities, not objective accuracy. A crucial next step is to investigate the relationship between Mint scores and physiological or behavioural measures of interoception (e.g., heartbeat tracking tasks, heartbeat evoked potentials). This would help establish the extent to which subjective reports on the Mint correspond to objective interoceptive performance. Future research could also systematically examine the cost of shorter versions, by investigating the tradeoffs incurred by versions with only the one or two most central facets per metacluster and their impact on psychometric qualities and predictive power.

While the overall structure of the Mint is stable, some facets warrant deeper investigation. The *Olfa* (Olfactory Compensation) and *Derm* (Dermal Hypersensitivity) clusters, for instance, showed more complex patterns of association. Their placement within the *Deficit* metacluster is theoretically intriguing but requires further validation. The former aligns with predictive coding models suggesting that individuals may rely on exteroceptive cues (like odours) to infer internal states  [@paulus2010interoception], a form of pseudo-interoception (or extero-interoception). The latter points to a potential link between visceral and cutaneous sensitivity, possibly reflecting a general trait of heightened bodily awareness or a shared mechanism of central sensitization, which has been observed in conditions like chronic pain [@di2016alexithymia]. Future studies should examine the invariance of the Mint's structure across different populations (e.g., clinical vs. non-clinical groups) and contexts (e.g., heightened arousal vs. rest) to clarify the role of these more nuanced facets. Such work could reveal whether their relevance and pattern of associations differs in conditions like anxiety disorders, chronic pain, or neurodevelopmental disorders.

In conclusion, the Mint scale represents a significant methodological advancement for the field of interoception research. By providing a validated, structurally coherent, and nuanced tool, it allows for a more precise assessment of self-reported interoceptive experiences. Its various forms (full, meta, mini, and micro) offer flexibility for researchers and clinicians, paving the way for a deeper understanding of the multifaceted role of interoception in human psychology and pathophysiology.

@@

We would like to thank the dissertation students from the University of Sussex for their help in data collection.
DM would also like to thank Anastasia for the motivation provided to write this paper.
